### 2023-08-22 introduction.tex
up

 \begin{center}
	\begin{figure}
	%		\hspace{3cm}
\includegraphics[width=0.45\textwidth]{FigureC}
\hspace{0.5cm}
\includegraphics[width=0.45\textwidth]{FigureD}
	\caption{Two reset-VASS with three states and one counter.}
					\label{r-V}
	\end{figure}
\end{center}

\begin{example}\label{Example}
{(\bf Reset-VASS example).}
Consider the leftmost reset-VASS from Figure~\ref{r-V}.
Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$. Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V$. However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of lenght less than $n$ towards $q_1(0)$. Consider now the rightmost reset-VASS, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$. In this reset-VASS { Bounded resilience} hold, since $10$-resilience hold. 
Indeed, from any configuration with value of the counter bigger than $10$, 
\end{example}

>>>>

 \begin{center}
	\begin{figure}
	%		\hspace{3cm}
\includegraphics[width=0.4\textwidth]{FigureC}
\hspace{0.6cm}
\includegraphics[width=0.4\textwidth]{FigureD}
	\caption{Two reset-VASS with three states and one counter.}
					\label{r-V}
	\end{figure}
\end{center}

\begin{example}\label{Example}
{(\bf Reset-VASS example).}
Consider the leftmost reset-VASS from Figure~\ref{r-V}.
Consider first $\Safe = \{q_0(n) \mid n \in \N\}$; In this case,  {Resilience}, { $k$-resilience} and { Bounded resilience} are not satisfied: there is no  way back from $q_2$ or $q_1$ to $q_0$. Consider now $\Safe = \{q_1(0)\} $; then { Resilience} hold: it is possible to reach $q_1(0)$ from any configuration of $V$. However, { Bounded resilience} do not hold, since, for all $n\in \N$, from $q_1(n)$, there is no path of lenght less than $n$ towards $q_1(0)$. Consider now the rightmost reset-VASS, which corresponds to having added a $+1$ transition from $q_1$ back to $q_0$. In this reset-VASS { Bounded resilience} hold, since $6$-resilience hold. 
Indeed, from any configuration with value of the counter bigger than $9$, it is possible to reach $q_2$ in two steps or less, then use the reset transition. From $q_1(n)$ or $q_0(n)$ with $n \in \{1, 2, 3, 4, 5, 6\}$ it is possible to reach $q_1(0)$ in $6$ steps or less. From 
$q_1(n)$ or $q_0(n)$ with $n \in \{7,8\}$ it is possible to reach $q_0(10)$ in $3$ steps or less, from which $q_1(0)$ is reachable in $2$ steps.
\end{example}

### 2023-08-22 resilience.tex
Qed
up

In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider two subsets of states $\Safe, \Bad \subseteq S$ such that $\Safe \cap  \Bad = \emptyset$ and $\Safe ,  \Bad \neq \emptyset$.
%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe,\Bad)$ is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp. a path of length smaller than or equal to $k$) that reaches a state in $\Safe$. We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp. $\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.  $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$). In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.

@@

In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
and $\Bad$ is the complement of $\Safe$. 
% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
% wheher
The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.

@@

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

\begin{proof}
Since $Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(Safe)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$.
		\end{proof}
%
%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.

>>>>

In a transition system $\mathscr{S}=(S,\rightarrow)$, we consider a subset of states $\Safe \subseteq S$, and its complement, $\Bad$.
%	The property of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
The \emph{resilience problem} (resp. the \emph{$k$-resilience problem}) for $(\mathscr{S},\Safe,\Bad)$ is to decide whether from \emph{any} state in $\Bad$, \emph{there exists} a path (resp. a path of length smaller than or equal to $k$) that reaches a state in $\Safe$. We use the notation $\Bad \longrightarrow^{*} \Safe$ (resp. $\Bad \longrightarrow^{\leq k} \Safe$) for $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{*} y$ (resp.  $\forall x \in \Bad, \exists y \in \Safe$ such that $x \longrightarrow^{\leq k} y$). In our framework, $\Safe, \Bad \subseteq S$  are possibly infinite but they must admit a computable finite representation : for example, downward-closed sets and upward-closed sets in wqo and semilinear sets in $\mathbb{N}^d$ have finite representations.

@@

\iffalse
In {\bf Flight Controller System}, the safe states correspond to the aircrafts having both landed, i.e.
$\Safe = \{(q_{A,1}, q_{M,i}, q_{B,1})(\textbf{v}) \mid \textbf{v} \in \N^4, i \in \N\}$,
and $\Bad$ is the complement of $\Safe$. 
% $\Bad = \{(q_{A,0}, q_{M,0}, q_{B,0})(\textbf{v}) \mid \textbf{v} \in \N^4\}$.
% wheher
The system is not $7$-resilient since there is no path of length smaller than $7$ from $(q_{A,0}, q_{M,5}, q_{B,0})(\textbf{v})$ to $\Safe$, but it is $8$-resilient.
\fi

@@

These three resilience problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS rely on the computation of upward or downward-closed sets \cite{DBLP:journals/iandc/AbdullaCJT00,DBLP:journals/tcs/FinkelS01}, we consider upward-closed or downward-closed sets $\Safe$ and $\Bad$. In \cite{DBLP:journals/corr/abs-2108-00889}, the authors considered that $\Bad$ is downward-closed and $\Safe$ is upward-closed. %\mathieu{Remarquer que dans l'example~\ref{Example} on a à la fois upward closed ET downward closed pour $\Safe$ ET $\Bad$.} 
% Remark that in {\bf Flight Controller System}, $\Safe$ and $\Bad$ are both upward-closed \emph{and} downward-closed.

@@

\begin{proof}
Since $Safe \subseteq S$ is upward-closed and $\mathscr{S}=(S, \rightarrow, \leq)$ is an effective WSTS with effective pred-basis, we compute the minimal basis of $\pred^*(Safe)=\{x_1,x_2,...,x_p\}$. 
%
%
Then we add the following guard $s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p$ to every transition $s \rightarrow s'$ of $\mathscr{S}$ and we define  $s \rightarrow' s'$ as $(s' \geq x_1  	\vee s' \geq x_2  	\vee...	\vee s' \geq x_p) \wedge s \rightarrow s'$ ; this insures that the new (guarded) WSTS $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is $\pred^*(Safe)$-resilient and then it is also the maximal resilient sub-WSTS of $\mathscr{S}=(S, \rightarrow, \leq)$. \qed
		\end{proof}
%
%It is also possible to define $(s_0,Safe)$-resilient systems as satisfying $\post^*(s_0) \xrightarrow{*} Safe$.
